Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-04-30 causal-bandit/nips_2016/introduction.tex
transfers icml paper to nips format

Our focus is on the simple regret, which measures the difference between the return of the optimal action and that of the action chosen by the algorithm after $T$ rounds.
In \S\ref{sec:simple-regret} we analyse a specific family of causal bandit problems that we call \emph{parallel bandit} problems in which $N$ factors affect the reward independently and there are $2N$ possible interventions.
We propose a simple causal best arm identification algorithm (Algorithm~\ref{alg:simple}) for this problem and show that up to logarithmic factors it enjoys minimax optimal
simple regret guarantees of $\tilde\Theta(\sqrt{m/T})$ where $m$ depends on the causal model and may be much smaller than $N$.
\todom{Need ref or proof?}
In contrast, existing best arm identification algorithms suffer $\Omega(\sqrt{N/T})$ simple regret.
This shows theoretically the value of our framework over the traditional bandit problem. 
Experiments in \S\ref{sec:experiments} further demonstrate the value of causal models in this framework.

@@

Two pieces of recent work also consider applying ideas from causal inference to bandit problems.
\citet{Bareinboim2015} demonstrate that in the presence of confounding variables the value that a variable would have taken had it not been intervened on can provide important contextual information. \todof{write one sentence to explain how their work is different to ours}

@@

>>>>

Our focus is on the simple regret, which measures the difference between the return of the optimal action and that of the action chosen by the algorithm after $T$ rounds.
In \S\ref{sec:simple-regret} we analyse a specific family of causal bandit problems that we call \emph{parallel bandit} problems in which $N$ factors affect the reward independently and there are $2N$ possible interventions.
We propose a simple causal best arm identification algorithm for this problem and show that up to logarithmic factors it enjoys minimax optimal
simple regret guarantees of $\tilde\Theta(\sqrt{m/T})$ where $m$ depends on the causal model and may be much smaller than $N$.
In contrast, existing best arm identification algorithms suffer $\Omega(\sqrt{N/T})$ simple regret (Thm. 4 by \citet{audibert2010best}).
This shows theoretically the value of our framework over the traditional bandit problem. 
Experiments in \S\ref{sec:experiments} further demonstrate the value of causal models in this framework.

@@

Two pieces of recent work also consider applying ideas from causal inference to bandit problems.
\citet{Bareinboim2015} demonstrate that in the presence of confounding variables the value that a variable would have taken had it not been 
intervened on can provide important contextual information. Their work differs in many ways. For example, the focus is on the cumulative regret and
the context is observed before the action is taken and cannot be controlled by the learning agent.

%\todof{write one sentence to explain how their work is different to ours}

@@

The truncated importance weighted estimators used in \S\ref{sec:simple-regret-general} have been studied before in a causal framework by \citet{BJQ13}, 
where the focus is on learning from observational data, but not controlling the sampling process. They also briefly discusses some of the issues 
encountered in sequential design, but do not give an algorithm or theoretical results for this case.

### 2016-04-30 causal-bandit/nips_2016/setup.tex
transfers icml paper to nips format

The objective of the learner is to minimise the simple regret
\begin{equation}
\label{eq:regret-simple}
	\simpleregret = \mu^* - \E{\mu_{\hat a^*_T}}.
\end{equation}
This is sometimes refered to as a ``pure exploration''~\citep{Bubeck2009} or ``best-arm identification'' problem~\citep{Gabillon2012} and
is most appropriate when, as in drug and policy testing, the learner has a fixed experimental budget after which its policy will be fixed indefintely.

>>>>

The objective of the learner is to minimise the simple regret
\begin{equation}
\label{eq:regret-simple}
	\simpleregret = \mu^* - \E{\mu_{\hat a^*_T}}.
\end{equation}
This is sometimes refered to as a ``pure exploration''~\citep{Bubeck2009a} or ``best-arm identification'' problem~\citep{Gabillon2012a} and
is most appropriate when, as in drug and policy testing, the learner has a fixed experimental budget after which its policy will be fixed indefintely.
